The History of Social Work Research: A Data Science Perspective

### 2015-01-29 Analysis.Rmd
Updated file to go with new bwr function.

# Data collection

+ Article record retrieval date:  January 8, 2015
Database: PsychInfo and Social Science Index (via EbscoHost Platform)

+ Search terms:  SO "social work" or SO "social welfare" or SO "social casework" or "social case work" or SO "social service" or "social services" or SO "human service" or SO "human services" or SO "social development" or SO "social environment"

@@

+ EbscoHost full journal listing retrieval date:  January 11, 2015 (http://www.ebscohost.com/title-lists)

# Data Pre-Processing
```{r initialization_dataprocessing, warning=FALSE, message=FALSE, eval=FALSE}
# Clear workspace
rm(list=ls())

# Read BWR functions for Mac OS
source("/Users/beperron/Git/BibWrangleR/functions/piWrangleR_Modified.R")

@@

# Wrangle the data with the BWR function suite
piBWR.f(csv=FALSE, path=path)
#save(pi.df, file = "piArticles.R")
```

```{r initialization_analysis, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
rm(list=ls())
setwd("/Users/beperron/Git/SocialWorkResearch/Data")
source("/Users/beperron/Git/BibWrangleR/functions/ggsurv.R")
load("ebscoFullListing.R")
load("piArticles.R")
load("socialWorkJournals.R")
library(dplyr)
library(ggplot2)
library(gridExtra)
library(survival)
library(grid)
library(png)
library(reshape2)

# Inspect dimensions of the data file (Rows X Columns)
dim(pi.df)

# Inspect variable names of the data file
names(pi.df)

# How many unique article titles?  Ebsco Results of most current search is $n=XXXX$. Do not proceed with analysis if the output does not match this result.  
length(which(pi.df$attributes == "TI"))
```

Journals excluded from search results.  _Social Work_ today excluded because it is of a magazine format.

```{r}
pi.df <- pi.df %>% filter(
        record != "PUT JOURNALS TO BE EXCLUDED HERE" &
        record != "PUT JOURNALS TO BE EXCLUDED HERE     " &
        )
```

Two journal titles were merged because one set of records was a logical abbreviation of another set of records.

```{r}

pi.df$record <- ifelse(pi.df$record == "Practice", "Practice: Social Work in Action"
                , pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 74, 42, pi.df$articleID)

pi.df$record <- ifelse(pi.df$record == "Eugenics & Social Welfare Bull.", "Eugenics and Social Welfare Bulletin", pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 28, 32, pi.df$articleID)
```

@@

All analyses provided here are not in publishable format, but presented to determine the basis of a storyline.  All tables and figures will be styled according to a target journal.

@@

```{r unique_titles, echo=FALSE}
unique.titles <- filter(pi.df, attributes == "SO")

@@

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, echo=FALSE}
journals.year <- tbl_df(pi.df)

@@

n.so.yr <- filter(pi.df, attributes == "SO" | attributes == "YR")

n.so <- filter(pi.df, attributes == "SO") %>% mutate(title = record) %>% 
        select(-attributes, -record)

n.yr <- filter(pi.df, attributes == "YR") %>% mutate(year = record ) %>% 
        select(-attributes, -record)

@@

# 10 highest number of publications
head(n.so.yr, 10)
```

@@

ggsurv(survival.journals) + 
    geom_line(data = median.survival, aes(time, quant), linetype="longdash") +
    annotate("segment", x = 32, xend = 27, y = .12, yend = .12, size = .25, arrow =arrow()) +
    geom_text(x = 45, y =.12, label = "median survival", size = 4) + 
    geom_text(x = 75, y =.75, label = "+ = Last year of indexing", size = 4) +
    ylim(0,1) +
    xlim(0,100) + 
    xlab("Survival Time (in years)") +
    ylab("Survival Probability") +
    ggtitle("Survival Probability of Social Work Journals")

@@

```{r articles_per_year, echo=FALSE}
n.articles.year <- filter(pi.df, attributes == "YR") 
year.split <- split(n.articles.year, n.articles.year$record)
year.count <- unlist(lapply(year.split, nrow))
year.count <- year.count[order(names(year.count))]
years <- names(year.count)

@@

```{r keywords_overall, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
kp.df <- filter(pi.df, attributes == "KP")

@@

```{r keywords_over_time, warning=FALSE, message=FALSE, echo=FALSE}
kp.yr <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

kp <- pi.df %>%
        filter(attributes == "KP") %>%
        select(articleID = articleID, keyword = record) %>%
        mutate(keyword = gsub(" ", "", keyword))

@@

```{r location, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
LO.df <- filter(pi.df, attributes == "LO") %>%
         group_by(record) %>%
         summarise(Country = unique(record),
                   N = n()) %>%
        arrange(desc(N)) %>% 
        select(-record)

@@

```{r location_over_time, warning=FALSE, message=FALSE, echo=FALSE}
library(RColorBrewer)
year <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

location.temp <- pi.df %>%
        filter(attributes == "LO") %>%
        select(articleID = articleID, country = record) %>%
        mutate(country = gsub(" ", "", country))

@@

## Number of authors
```{r comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
n.authors.article <- pi.df %>% 
    filter(attributes == "AU") %>%
    select(id = articleID, author= record) %>%
    mutate(id = as.numeric(id))

@@

```{r message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4, echo=FALSE}
df.2 <- tbl_df(pi.df)
year <- df.2 %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)

>>>>

# Data harvesting

+ Article record retrieval date: January 24, 2015
Database: psycInfo, Social Science Abstracts, Social Work Abstracts (via EbscoHost Platform)

+ Primary search:  Hodge's list

+ Supplemental search:  SO "social work" or SO "social welfare" or SO "social casework" or "social case work" or SO "social service" or "social services" or SO "human service" or SO "human services" or SO "social development" or SO "social environment"

@@

+ Results filtered to exclude years 2014 and 2015 (due to delays in indexing)

# Data Pre-Processing

```{r initialization_dataprocessing, warning=FALSE, message=FALSE, eval=FALSE}
 # Clear workspace
rm(list=ls())

# Read BWR functions for Mac OS
source("/Users/beperron/Git/BibWrangleR/functions/ebscoBWR.R")

@@

# Wrangle the data with the BWR function suite
ebscoBWR.f(csv=FALSE, path=path)
#save(bwr.df, file = "piArticles.R")
```

```{r initialization_analysis, message=FALSE, warning=FALSE, comment=NA}
rm(list=ls())
setwd("/Users/beperron/Git/SocialWorkResearch/Data")
source("/Users/beperron/Git/BibWrangleR/functions/ggsurv.R")

library(dplyr)
library(ggplot2)
library(gridExtra)
library(survival)
library(grid)
library(png)
library(reshape2)

# Inspect dimensions of the data file (Rows X Columns)
dim(bwr.df)

# Inspect variable names of the data file
names(bwr.df)

# How many unique article titles?  Ebsco Results of most current search is $n=XXXX$. Do not proceed with analysis if the output does not match this result.  
length(which(bwr.df$attributes == "TI"))
```

# Exclude years 2014 and 2015

```{r}
years <- c("2014", "2015")
ebsco.remove <- bwr.df[bwr.df$record %in% years, ]
ebsco.remove <- ebsco.remove[, "articleID"]
bwr.df <- bwr.df[!(bwr.df$articleID %in% ebsco.remove),]
```

# Merge journal titles

```{r}
bwr.df$record <- ifelse(bwr.df$record == "Eugenics and Social Welfare Bulletin", 
                       "Eugenics & Social Welfare Bull.", 
                        bwr.df$record)

bwr.df$record <- ifelse(bwr.df$record == "The British Journal of Social Work", 
                       "British Journal of Social Work", 
                        bwr.df$record)

bwr.df$record <- ifelse(bwr.df$record == "Social Work & Social Sciences Review", 
                       "Social Work and Social Sciences Review", 
                        bwr.df$record)

bwr.df$record <- ifelse(bwr.df$record == "Social Work with Groups", 
                       "Social Work With Groups", 
                        bwr.df$record)

bwr.df$record <- ifelse(bwr.df$record == "Computers in Human Services", 
                       "Journal of Technology in the Human Services", 
                        bwr.df$record)

# Journals excluded from search results.

```{r}
exclusions <- c(
    "American Journal of Preventive Medicine", 
    "Behavior Modification", 
    "Brain and Cognition",
    "Canadian Journal of Community Mental Health",
    "Canadian Journal on Aging",
    "Early Child Development and Care",
    "Early Education and Development",
    "Employee Assistance Quarterly",
    "Evaluation and Program Planning",
    "General Hospital Psychiatry",
    "Hospital Social Service",
    "Journal of Applied Behavioral Science",
    "Journal of Applied Rehabilitation Counseling",
    "Journal of Community & Applied Social Psychology",
    "Journal of Nonverbal Behavior",
    "Learning Disabilities Research & Practice",
    "New York State Department of Social Welfare, Division Publication",
    "Pennsylvania Social Work",
    "Proceedings of the National Converence of Social Work",
    "PROFILE Issues in Teachers' Professional Development",
    "Psychotherapy",
    "Social Development",
    "Social Service Department, Presbyterian Hospital, New York",
    "Social Work Technique",
    "Social Work Today", # Magazine
    "The New Social Worker",
    "Systems Research and Behavioral Science",
    "Teaching and Teacher Education",
    "University of Chicago Social Service Monograph",
    "The Scientific Review of Mental Health Practice",
    "The Clinical Supervisor",
    "University of Chicago Social Service Monograph")
ebsco.remove <- bwr.df[bwr.df$record %in% exclusions, ]
ebsco.remove <- ebsco.remove[, "articleID"]
bwr.df <- bwr.df[!(bwr.df$articleID %in% ebsco.remove),]
```

filter(bwr.df, attributes == "SO") %>% group_by(record) %>% arrange(record) %>% summarise(Article.Count = n())

filter(bwr.df, attributes == "TI") %>% summarize(N = n())
```

@@

@@

```{r unique_titles, echo=FALSE}
unique.titles <- filter(bwr.df, attributes == "SO")

@@

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, echo=FALSE}
journals.year <- tbl_df(bwr.df)

@@

n.so.yr <- filter(bwr.df, attributes == "SO" | attributes == "YR")

n.so <- filter(bwr.df, attributes == "SO") %>% mutate(title = record) %>% 
        select(-attributes, -record)

n.yr <- filter(bwr.df, attributes == "YR") %>% mutate(year = record ) %>% 
        select(-attributes, -record)

@@

# 10 highest number of publications
head(n.so.yr, 20)
```

@@

ggsurv(survival.journals) + 
    #geom_line(data = median.survival, aes(time, quant), linetype="longdash") +
    #annotate("segment", x = 32, xend = 27, y = .12, yend = .12, size = .25, arrow =arrow()) +
    #geom_text(x = 45, y =.12, label = "median survival", size = 4) + 
    geom_text(x = 75, y =.75, label = "+ = Last year of indexing", size = 4) +
    ylim(0,1) +
    xlim(0,50) + 
    xlab("Survival Time (in years)") +
    ylab("Survival Probability") +
    ggtitle("Survival Probability of Social Work Journals")

@@

```{r articles_per_year, echo=FALSE}
n.articles.year <- filter(bwr.df, attributes == "YR") 
year.split <- split(n.articles.year, n.articles.year$record)
year.count <- unlist(lapply(year.split, nrow))
year.count <- year.count[order(names(year.count))]
years <- names(year.count)

@@

```{r keywords_overall, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
kp.df <- filter(bwr.df, attributes == "KP")

@@

```{r keywords_over_time, warning=FALSE, message=FALSE, echo=FALSE}
kp.yr <- filter(bwr.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

kp <- bwr.df %>%
        filter(attributes == "KP") %>%
        select(articleID = articleID, keyword = record) %>%
        mutate(keyword = gsub(" ", "", keyword))

@@

```{r location, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
LO.df <- filter(bwr.df, attributes == "LO") %>%
         group_by(record) %>%
         summarise(Country = unique(record),
                   N = n()) %>%
        arrange(desc(N)) %>% 
        select(-record)

@@

```{r location_over_time, warning=FALSE, message=FALSE, echo=FALSE}
library(RColorBrewer)
year <- filter(bwr.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

location.temp <- bwr.df %>%
        filter(attributes == "LO") %>%
        select(articleID = articleID, country = record) %>%
        mutate(country = gsub(" ", "", country))

@@

## Number of authors
```{r comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
n.authors.article <- bwr.df %>% 
    filter(attributes == "AU") %>%
    select(id = articleID, author= record) %>%
    mutate(id = as.numeric(id))

@@

```{r message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4, echo=FALSE}
df.2 <- tbl_df(bwr.df)
year <- df.2 %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)
